---
title: "Baseline Dataset for paper on Deaf Depression-free Life Expectancy"
author: "Rafael Rodrigues de Moraes"
date: "2024-11-10"
output:
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Packages

The following packages are needed to load the datasets (`arrow` and
`readr`), applyt the complex survey design (`survey`) to the 
National Health Survey (PNS2019) selected microdata and
prepare the data frames (`dplyr`, `tidyr`) for the analyses in the
paper.

```{r packages, message=FALSE, warning=FALSE, include=TRUE}
library(tidyverse)
library(arrow)     # read_parquet
library(survey)    # svydesign, postStratify, svytable
```

# [`br_ibge_pns_2019`] Importing the processed National Health Survey 2019 dataset

```{r br_ibge_pns_2019, include=TRUE}
# 2019th edition from National Health Survey (PNS/IBGE)
br_ibge_pns_2019 <- 
  read_parquet(
    file = '/home/rstudio/data/br_ibge_pns_2019/microdata.parquet'
  ) %>% mutate_at(c('v0028','v00282','v00283'), as.numeric)
```

# Complex Survey Design

The National Health Survey (PNS) is a complex survey consisting of
strata, cluster and weights.

The same steps were followed in order to set up the survey design object
and post-stratify it as listed in the official R package `PNSIBGE`
(version 0.1.7).

```{r pns_survey_design, include=TRUE}
## Survey design (source: https://rdrr.io/cran/PNSIBGE/src/R/pns_design.R, accessed on September, 2nd 2022)
options(survey.lonely.psu="adjust")
options(survey.adjust.domain.lonely=TRUE)

pns2019_prior <- survey::svydesign(
  ids = ~upa_pns, 
  strata = ~v0024, 
  data = subset(br_ibge_pns_2019, !is.na(v0028)), 
  weights = ~v0028, 
  nest = TRUE
)
# Post-stratification
popc.types <- data.frame( 
  v00283=as.character(unique(br_ibge_pns_2019$v00283)), 
  Freq=as.numeric(unique(br_ibge_pns_2019$v00282))
)
popc.types <- popc.types[order(popc.types$v00283),]
pns2019_posterior <- survey::postStratify(
  design=pns2019_prior, 
  strata=~v00283, 
  population=popc.types
)
```

# Baseline Dataset for Deaf Depression-free Life Expectancy Paper

```{r dfle_baseline_preparation, include=TRUE}
dfle_raw <- svytable(
  formula=~c008 + # age at interview date
    g058 +        # level of hearing impairment
    g057 +        # level of hearing impairment (despite use of hearing devices)
    c006 +        # sex 
    g05801 +      # use of LIBRAS, the Brazilian sign language
    q092 +        # ever diagnosed with depression by a physician
    q11006,       # ever diagnosed with anxiety by a physician
  design = pns2019_posterior
)
```

```{r dfle_baseline_export, include=TRUE}
saveRDS(
  dfle_raw, 
  file = '/home/rstudio/data/dfle_baseline_dataset.rds'
)
```
