---
title: "Figure01 -- Paper on Deaf Depression-free Life Expectancy"
author: "Rafael Rodrigues de Moraes"
date: "2024-11-10"
output:
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning = FALSE)
```

# R Packages

The following packages are needed to load the datasets (`arrow` and
`readr`), applyt the complex survey design (`survey`) to the 
National Health Survey (PNS2019) selected microdata and
prepare the data frames (`dplyr`, `tidyr`) for the analyses in the
paper.

```{r libraries, message=FALSE, warning=FALSE, include=TRUE}
library(readr)     # read_delim
library(tidyr)     # pivot_wider
library(dplyr)     # %>%, filter, select, 
library(ggplot2)   # geom_point, ggplot
```

# Importing the baseline dataset
```{r baseline_dataset_import, include=TRUE}
dfle_raw <-
  readRDS(file = '/home/rstudio/data/dfle_baseline_dataset.rds')
```

## Figure 1: Depression prevalence by sex and age band among grouped levels of hearing impairment, Brazil, 2019

```{r dfle_figure01, include=TRUE}
AGE_BREAKS <- c( 0, 1, seq(from=5,to=80, by=5), 1e3)
AGE_LABELS <- c('<1','1-4','5-9','10-14','15-19','20-24','25-29','30-34'
                ,'35-39','40-44','45-49','50-54','55-59','60-64','65-69'
                ,'70-74','75-79','80+')

dfle_pns <- 
  dfle_raw %>% 
  as.data.frame(stringsAsFactors = FALSE) %>%
  mutate(
    c008 = as.numeric(c008),
    sex = factor( c006, levels = 1:2, labels = c('male', 'female')),
    libras = factor( g05801, levels = 1:2, labels = c('yes', 'no')),
    age_grp = cut( 
      c008, 
      breaks = AGE_BREAKS,
      labels = AGE_LABELS,
      include.lowest = TRUE,
      right = FALSE
    ),
    combined = case_when(
      g057 == ' ' ~ g058,
      g058 == ' ' ~ g057
    ),
    hearing_impairment_level = case_when(
      combined %in% c('3', '4') ~ 'Heavily or Fully impaired',
      combined %in% c('1', '2') ~ 'Not or mildly impaired',
      # TRUE ~ 'Ignored or missing information'
    ),
    hearing_impaired = case_when(
      combined == '4' ~ 'Fully impaired',
      combined == '3' ~ 'Heavily impaired',
      combined == '2' ~ 'Mildly impaired',
      combined == '1' ~ 'Not impaired',
      # TRUE ~ 'Ignored or missing information'
    ),
    diagnosed_depression = q092,
    diagnosed_anxiety = q11006,
)

# ever diagnosed with depression by a physician (g058)
depr_by_impairment <- 
  dfle_pns %>%
  na.omit() %>%
  group_by( hearing_impairment_level, age_grp, sex, diagnosed_depression) %>%
  summarise( n = sum( Freq ) ) %>%
  filter( diagnosed_depression != ' ') %>%
  pivot_wider( id_cols = c(1,2,3), names_from = 4, values_from = n) %>% 
  rename( yes = `1`, no = `2`) 

depr_by_impairment <- 
  bind_rows(
    depr_by_impairment,
    depr_by_impairment %>% 
    group_by( hearing_impairment_level, age_grp) %>%
    summarise( yes = sum(yes), no = sum(no)) %>%
    transmute( hearing_impairment_level, age_grp, sex = 'both', yes, no)
  ) %>% 
  mutate(
    N = (yes + no),
    pct = ifelse( yes+no== 0, NA, yes / N),
    var_pct = ifelse(N==0, NA, ( pct * (1-pct) )/(N))
  ) 

fig1 <-
  depr_by_impairment %>%
  filter(!age_grp %in% c('<1','1-4','5-9','10-14')) %>%
  ggplot(
    aes(
      x = age_grp,
      y = pct,
      group = sex,
      colour = hearing_impairment_level
    )
  ) +
  geom_errorbar(
    aes(
      ymin = pct - 1.96 * sqrt(var_pct), 
      ymax = pct + 1.96 * sqrt(var_pct)
    ),
    width = 0.5 
  ) +
  geom_point(size = 1.5, alpha = .8) +
  labs(
    y = 'People diagnosed with depression by a physician'
    ,x = 'Age groups'
    ,group = ''
    ,colour = ''
  ) +
  theme_bw(base_size = 13) +
  theme(
    panel.grid = element_blank(),
    legend.position = 'top'
  )+
  facet_grid(~sex) +
  coord_flip() +
  scale_color_brewer(palette = "Dark2") +
  scale_y_continuous(
    labels = scales::percent_format(scale = 100)
  )


for( ext in c('png','pdf','svg')){
  ggsave(
    filename = paste0('fig01.',ext),
    plot = fig1,
    width = 7,
    height = 5
  )  
}

fig1
```

