---
title: "Support analyses for paper on Deaf Depression-free Life Expectancy"
author: "Rafael Rodrigues de Moraes"
date: "2024-11-10"
output:
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

This RMarkdown notebook serves as the official documentation of the
steps undertaken to produce the data and graphics available on
the Deaf Depression-free Life Expectancy paper.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Packages

The following packages are needed to load the datasets (`arrow` and
`readr`), applyt the complex survey design (`survey`) to the 
National Health Survey (PNS2019) selected microdata and
prepare the data frames (`dplyr`, `tidyr`) for the analyses in the
paper.

```{r libraries, message=FALSE, warning=FALSE, include=TRUE}
library(tidyverse)
library(ggstats)   # 
library(svglite)   # save plots to SVG via ggsave
library(arrow)     # read_parquet
library(survey)    # svydesign, postStratify, svytable
```

# Importing the processed datasets

## [`br_ibge_censo_2010`] Abridged Life Tables from brazilian census (edition 2010)

```{r br_ibge_censo_2010, include=TRUE}
# Life tables from 2010th Brazilian Population Census (IBGE)
br_igbe_censo_2010 <-
  read_delim(
    file = '/home/rstudio/data/br_ibge_censo_2010/censo2010_lifetables.csv',
    delim = '|',
    show_col_types = FALSE
  )
```

## [`br_ibge_pns_2019`] National Health Survey 2019

```{r br_ibge_pns_2019, include=TRUE}
# 2019th edition from National Health Survey (PNS/IBGE)
br_ibge_pns_2019 <- read_parquet(
  file = '/home/rstudio/data/br_ibge_pns_2019/microdata.parquet'
) %>%
  mutate(across(everything(), as.numeric))
```

# Complex Survey Design

The National Health Survey (PNS) is a complex survey consisting of
strata, cluster and weights.

The same steps were followed in order to set up the survey design object
and post-stratify it as listed in the official R package `PNSIBGE`
(version 0.1.7).

```{r pns_survey_design, include=TRUE}
## Survey design (source: https://rdrr.io/cran/PNSIBGE/src/R/pns_design.R, accessed on September, 2nd 2022)
options(survey.lonely.psu="adjust")
options(survey.adjust.domain.lonely=TRUE)

pns2019_prior <- survey::svydesign(
  ids = ~upa_pns, 
  strata = ~v0024, 
  data = subset(br_ibge_pns_2019, !is.na(v0028)), 
  weights = ~v0028, 
  nest = TRUE
)
# Post-stratification
popc.types <- data.frame( 
  v00283=as.character(unique(br_ibge_pns_2019$v00283)), 
  Freq=as.numeric(unique(br_ibge_pns_2019$v00282))
)
popc.types <- popc.types[order(popc.types$v00283),]
pns2019_posterior <- survey::postStratify(
  design=pns2019_prior, 
  strata=~v00283, 
  population=popc.types
)
```

## Validation

Despite using exactly the same complex survey design, a full match with
the government official statistics was not entirely possible.

![PNS2019 - Validation with self-reported health
perception](./pns_validation.png){width="70%"}

```{r pns_validation, include=TRUE}
# validation against
#   https://sidra.ibge.gov.br/tabela/7666#/n1/all/v/2667/p/all/c2/all/c12258/all/d/v2667%200/l/v,p+c2,t+c12258/resultado
round( svytable(formula=~j001+c006, design=subset(pns2019_posterior, c008 >=18))/1e3, 0) %>%
  as.data.frame() %>%
  mutate(
    tab7666 = case_when(
      j001 %in% c(1,2) ~ 1,
      j001 == 3 ~ 2,
      j001 %in% c(4,5) ~ 3,
    )
    ,tab7666 = factor( tab7666, levels = 1:3, labels = c('Muito ruim e ruim','Regular','Muito bom e bom'))
    ,c006 = factor( c006, levels = 1:2, labels = c('Masculino', 'Feminino'))
  ) %>%
  na.omit() %>%
  group_by(tab7666, c006) %>%
  summarise( n = sum(Freq) ) %>%
  pivot_wider(id_cols = tab7666, names_from = c006, values_from = n)
```

```{r pns_validation_1, include=TRUE}
round( svytable(formula=~c006, design=subset(pns2019_posterior, c008 >= 18))/1e3, 0)
```

```{r pns_validation_2, include=TRUE}
round( sum(svytable(formula=~c006, design=subset(pns2019_posterior, c008 >= 18)))/1e3, 0)
```

*Conclusion:* although overall total and totals by sex are OK, among
health perception the numbers do not match. Unfortunately it was not
possible to reconcile the differences but as the total figures are the
same, we proceed with the analysis.

# Deaf Depression-free Life Expectancy paper related analyses

```{r dfle_data_preparation, include=TRUE}
dfle_raw <- svytable(
  formula=~c008 + # age at interview date
    g058 +        # level of hearing impairment
    g057 +        # level of hearing impairment (despite use of hearing devices)
    c006 +        # sex 
    g05801 +      # use of LIBRAS, the Brazilian sign language
    q092 +        # ever diagnosed with depression by a physician
    q11006,       # ever diagnosed with anxiety by a physician
  design = pns2019_posterior,
)
```

## Table 2: Libras users and prevalence of depression for respondents with age between 5 and 40 years, Brazil, 2019

```{r dfle_table02, include=TRUE}
AGE_BREAKS <- c( 0, 1, seq(from=5,to=80, by=5), 1e3)
AGE_LABELS <- c('<1','1-4','5-9','10-14','15-19','20-24','25-29','30-34'
                ,'35-39','40-44','45-49','50-54','55-59','60-64','65-69'
                ,'70-74','75-79','80+')

dfle_pns <- dfle_raw %>% 
  as.data.frame(stringsAsFactors = FALSE) %>%
  mutate(
    c008 = as.numeric(c008),
    sex = factor( c006, levels = 1:2, labels = c('male', 'female')),
    libras = factor( g05801, levels = 1:2, labels = c('yes', 'no')),
    age_grp = cut( 
      c008, 
      breaks = AGE_BREAKS,
      labels = AGE_LABELS,
      include.lowest = TRUE,
      right = FALSE
    ),
    combined = case_when(
      g057 == ' ' ~ g058,
      g058 == ' ' ~ g057,
    ),
    hearing_impairment_level = case_when(
      combined %in% c('3', '4') ~ 'Heavily or Fully impaired',
      combined %in% c('1', '2') ~ 'Not or mildly impaired',
      # TRUE ~ 'Ignored or missing information'
    ),
    hearing_impaired = case_when(
      combined == '4' ~ 'Fully impaired',
      combined == '3' ~ 'Heavily impaired',
      combined == '2' ~ 'Mildly impaired',
      combined == '1' ~ 'Not impaired',
      # TRUE ~ 'Ignored or missing information'
    ),
    diagnosed_depression = q092,
    diagnosed_anxiety = q11006,
  ) 

summary_aux <- dfle_pns %>% 
  filter( c008 >= 5, c008 <= 40 ) %>%
  group_by(combined, hearing_impaired, libras, diagnosed_depression) %>%
  summarise( n = sum(Freq))
  
summary_aux %>%
  group_by(combined, hearing_impaired, libras) %>%
  summarise( n = sum(n)) %>%
  na.omit() %>%
  pivot_wider(id_cols = c(1,2), names_from = 3, values_from=4) %>%
  mutate( pct_libras_users = yes / (yes+no)) %>%
  select( -c(yes,no)) %>%
  left_join( summary_aux %>% 
               group_by(combined, hearing_impaired, diagnosed_depression) %>% 
               summarise( n = sum(n)) %>%
               na.omit() %>% 
               pivot_wider(id_cols = c(1,2), names_from = 3, values_from=4) %>% 
               mutate( pct_depression = `1` / (`1`+`2`)) %>% 
               select( -c(3:5)) 
             ) %>%
  left_join( summary_aux %>% 
               filter( libras == 'yes') %>% 
               group_by(combined, hearing_impaired, diagnosed_depression) %>% 
               summarise( n = sum(n)) %>% 
               na.omit() %>% 
               pivot_wider(id_cols = c(1,2), names_from = 3, values_from=4) %>% 
               mutate( pct_depression_among_libras_users = `1` / (`1`+`2`)) %>%
               select( -c(3:5)) 
             ) %>% 
  ungroup() %>%
  select( -combined ) %>%
  knitr::kable()
```

## Figure 1: Depression prevalence by sex and age band among grouped levels of hearing impairment, Brazil, 2019

```{r dfle_figure01, include=TRUE}
AGE_BREAKS <- c( 0, 1, seq(from=5,to=80, by=5), 1e3)
AGE_LABELS <- c('<1','1-4','5-9','10-14','15-19','20-24','25-29','30-34'
                ,'35-39','40-44','45-49','50-54','55-59','60-64','65-69'
                ,'70-74','75-79','80+')

dfle_pns <- dfle_raw %>% 
  as.data.frame(stringsAsFactors = FALSE) %>%
  mutate(
    c008 = as.numeric(c008),
    sex = factor( c006, levels = 1:2, labels = c('male', 'female')),
    libras = factor( g05801, levels = 1:2, labels = c('yes', 'no')),
    age_grp = cut( 
      c008, 
      breaks = AGE_BREAKS,
      labels = AGE_LABELS,
      include.lowest = TRUE,
      right = FALSE
    ),
    combined = case_when(
      g057 == ' ' ~ g058,
      g058 == ' ' ~ g057,
    ),
    hearing_impairment_level = case_when(
      combined %in% c('3', '4') ~ 'Heavily or Fully impaired',
      combined %in% c('1', '2') ~ 'Not or mildly impaired',
      # TRUE ~ 'Ignored or missing information'
    ),
    hearing_impaired = case_when(
      combined == '4' ~ 'Fully impaired',
      combined == '3' ~ 'Heavily impaired',
      combined == '2' ~ 'Mildly impaired',
      combined == '1' ~ 'Not impaired',
      # TRUE ~ 'Ignored or missing information'
    ),
    diagnosed_depression = q092,
    diagnosed_anxiety = q11006,
  ) 

# ever diagnosed with depression by a physician (g058)
depr_by_impairment <- dfle_pns %>%
  na.omit() %>%
  group_by( hearing_impairment_level, age_grp, sex, diagnosed_depression) %>%
  summarise( n = sum( Freq ) ) %>%
  filter( diagnosed_depression != ' ') %>%
  pivot_wider( id_cols = c(1,2,3), names_from = 4, values_from = n) %>% 
  rename( yes = `1`, no = `2`) 

depr_by_impairment <- bind_rows(
  depr_by_impairment,
  depr_by_impairment %>% 
  group_by( hearing_impairment_level, age_grp) %>%
  summarise( yes = sum(yes), no = sum(no)) %>%
  transmute( hearing_impairment_level, age_grp, sex = 'both', yes, no)
) %>% mutate(
    pct = ifelse( yes+no== 0, 0, yes / (yes + no))
  ) 

fig1 <- depr_by_impairment %>%
  filter(!age_grp %in% c('<1','1-4','5-9','10-14')) %>%
  ggplot(
    aes(
      x = age_grp,
      y = pct,
      group = sex,
      colour = hearing_impairment_level
    )
  ) +
  geom_line(
    aes(group = age_grp),
    size = 1,
    color = "grey26",
    alpha = 1
    ) +
  geom_point(size = 4.3, alpha = .8) +
  geom_stripped_cols(color = "white") +
  labs(
    y = 'People diagnosed with depression by a physician'
    ,x = 'Age groups'
    ,group = ''
    ,colour = ''
  ) +
  theme_bw(base_size = 13) +
  theme(
    panel.grid = element_blank(),
    legend.position = 'top'
  )+
  facet_grid(~sex) +
  coord_flip() +
  scale_color_brewer(palette = "Dark2") +
  scale_y_continuous(
    labels = scales::percent_format(scale = 100)
    )

ggsave(
  filename = 'fig01.png',
  plot = fig1,
  width = 7,
  height = 5
)

ggsave(
  filename = 'fig01.pdf',
  plot = fig1,
  width = 7,
  height = 5
)

ggsave(
  filename = 'fig01.svg',
  plot = fig1,
  width = 7,
  height = 5
)

fig1
```

## Table 3: Comparison by sex, age, and grouped level of hearing impairment of life expectancy and depression-free life expectancy, Brazil

```{r dfle_table03, include=TRUE}
AGE_BREAKS <- c( 0, 1, seq(from=5,to=90, by=5), 1e3)
AGE_LABELS <- c('<1','1-4','5-9','10-14','15-19','20-24','25-29','30-34'
                ,'35-39','40-44','45-49','50-54','55-59','60-64','65-69'
                ,'70-74','75-79','80-84','85-89','90+')

dfle_pns <- dfle_raw %>% 
  as.data.frame(stringsAsFactors = FALSE) %>%
  mutate(
    c008 = as.numeric(c008),
    sex = factor( c006, levels = 1:2, labels = c('male', 'female')),
    libras = factor( g05801, levels = 1:2, labels = c('yes', 'no')),
    age_grp = cut( 
      c008, 
      breaks = AGE_BREAKS,
      labels = AGE_LABELS,
      include.lowest = TRUE,
      right = FALSE
    ),
    combined = case_when(
      g057 == ' ' ~ g058,
      g058 == ' ' ~ g057,
    ),
    hearing_impairment_level = case_when(
      combined %in% c('3', '4') ~ 'Heavily or Fully impaired',
      combined %in% c('1', '2') ~ 'Not or mildly impaired',
      # TRUE ~ 'Ignored or missing information'
    ),
    hearing_impaired = case_when(
      combined == '4' ~ 'Fully impaired',
      combined == '3' ~ 'Heavily impaired',
      combined == '2' ~ 'Mildly impaired',
      combined == '1' ~ 'Not impaired',
      # TRUE ~ 'Ignored or missing information'
    ),
    diagnosed_depression = q092,
    diagnosed_anxiety = q11006,
  ) 

# ever diagnosed with depression by a physician (g058)
depr_by_impairment <- dfle_pns %>%
  na.omit() %>%
  group_by( hearing_impairment_level, age_grp, sex, diagnosed_depression) %>%
  summarise( n = sum( Freq ) ) %>%
  filter( diagnosed_depression != ' ') %>%
  pivot_wider( id_cols = c(1,2,3), names_from = 4, values_from = n) %>% 
  rename( yes = `1`, no = `2`) 

depr_by_impairment <- bind_rows(
  depr_by_impairment,
  depr_by_impairment %>% 
  group_by( hearing_impairment_level, age_grp) %>%
  summarise( yes = sum(yes), no = sum(no)) %>%
  transmute( hearing_impairment_level, age_grp, sex = 'both', yes, no)
) %>% mutate(
    pct = ifelse( yes+no== 0, 0, yes / (yes + no))
  ) 

# ever diagnosed with anxiety by a physician (q11006)
anx_by_impairment <- dfle_pns %>%
  na.omit() %>%
  group_by( hearing_impairment_level, age_grp, sex, diagnosed_anxiety) %>%
  summarise( n = sum( Freq ) ) %>%
  filter( diagnosed_anxiety != ' ') %>%
  pivot_wider( id_cols = c(1,2,3), names_from = 4, values_from = n) %>% 
  rename( yes = `1`, no = `2`) 

anx_by_impairment <- bind_rows(
  anx_by_impairment,
  anx_by_impairment %>% 
    group_by( hearing_impairment_level, age_grp) %>%
    summarise( yes = sum(yes), no = sum(no)) %>%
    transmute( hearing_impairment_level, age_grp, sex = 'both', yes, no)
) %>% mutate(
  pct = ifelse( yes+no== 0, 0, yes / (yes + no))
) 

prevalences <- bind_rows(
  depr_by_impairment %>%
  mutate( disease = 'depression'),
  anx_by_impairment %>%
  mutate( disease = 'anxiety') ) 

dfle <- prevalences %>% 
  transmute(disease, sex, hearing_impairment_level, age_grp, pix = pct) %>%
  arrange( disease, sex,hearing_impairment_level, desc(age_grp) ) %>% 
  na.omit() %>%
  left_join(br_ibge_censo_2010, by = c('age_grp','sex')) %>% 
  mutate( aux = (1-pix)*Lxn ) %>% 
  group_by(disease, sex, hearing_impairment_level) %>%
  mutate( 
    dfle = cumsum(aux)/lx, 
    age_grp = factor( age_grp, levels = AGE_LABELS),
    dfle_over_Ex = dfle / Ex
    ) %>% 
  select( -aux ) %>% 
  arrange( disease, sex, hearing_impairment_level, age_grp ) 

dfle %>% 
  filter(
    disease == 'depression',
    age_grp %in% c('<1', '20-24', '40-44','60-64','80-84')
  ) %>%
  mutate(
    age = case_when(
      age_grp == '<1' ~ 0,
      age_grp == '20-24' ~ 20,
      age_grp == '40-44' ~ 40,
      age_grp == '60-64' ~ 60,
      age_grp == '80-84' ~ 80,
    )
  ) %>%
  pivot_wider(
    id_cols = c('sex', 'age', 'Ex'),
    names_from = 'hearing_impairment_level',
    values_from = 'dfle'
  ) %>%
  select( 1,2,3,5,4) %>%
  knitr::kable()
```


## Figure 2: Proportional impact of depression on life expectancy according to sex and grouped hearing impairment level, Brazil

```{r dfle_figure02, include=TRUE}
AGE_BREAKS <- c( 0, 1, seq(from=5,to=90, by=5), 1e3)
AGE_LABELS <- c('<1','1-4','5-9','10-14','15-19','20-24','25-29','30-34'
                ,'35-39','40-44','45-49','50-54','55-59','60-64','65-69'
                ,'70-74','75-79','80-84','85-89','90+')

dfle_pns <- dfle_raw %>% 
  as.data.frame(stringsAsFactors = FALSE) %>%
  mutate(
    c008 = as.numeric(c008),
    sex = factor( c006, levels = 1:2, labels = c('male', 'female')),
    libras = factor( g05801, levels = 1:2, labels = c('yes', 'no')),
    age_grp = cut( 
      c008, 
      breaks = AGE_BREAKS,
      labels = AGE_LABELS,
      include.lowest = TRUE,
      right = FALSE
    ),
    combined = case_when(
      g057 == ' ' ~ g058,
      g058 == ' ' ~ g057,
    ),
    hearing_impairment_level = case_when(
      combined %in% c('3', '4') ~ 'Heavily or Fully impaired',
      combined %in% c('1', '2') ~ 'Not or mildly impaired',
      # TRUE ~ 'Ignored or missing information'
    ),
    hearing_impaired = case_when(
      combined == '4' ~ 'Fully impaired',
      combined == '3' ~ 'Heavily impaired',
      combined == '2' ~ 'Mildly impaired',
      combined == '1' ~ 'Not impaired',
      # TRUE ~ 'Ignored or missing information'
    ),
    diagnosed_depression = q092,
    diagnosed_anxiety = q11006,
  ) 

# ever diagnosed with depression by a physician (g058)
depr_by_impairment <- dfle_pns %>%
  na.omit() %>%
  group_by( hearing_impairment_level, age_grp, sex, diagnosed_depression) %>%
  summarise( n = sum( Freq ) ) %>%
  filter( diagnosed_depression != ' ') %>%
  pivot_wider( id_cols = c(1,2,3), names_from = 4, values_from = n) %>% 
  rename( yes = `1`, no = `2`) 

depr_by_impairment <- bind_rows(
  depr_by_impairment,
  depr_by_impairment %>% 
  group_by( hearing_impairment_level, age_grp) %>%
  summarise( yes = sum(yes), no = sum(no)) %>%
  transmute( hearing_impairment_level, age_grp, sex = 'both', yes, no)
) %>% mutate(
    pct = ifelse( yes+no== 0, 0, yes / (yes + no))
  ) 

# ever diagnosed with anxiety by a physician (q11006)
anx_by_impairment <- dfle_pns %>%
  na.omit() %>%
  group_by( hearing_impairment_level, age_grp, sex, diagnosed_anxiety) %>%
  summarise( n = sum( Freq ) ) %>%
  filter( diagnosed_anxiety != ' ') %>%
  pivot_wider( id_cols = c(1,2,3), names_from = 4, values_from = n) %>% 
  rename( yes = `1`, no = `2`) 

anx_by_impairment <- bind_rows(
  anx_by_impairment,
  anx_by_impairment %>% 
    group_by( hearing_impairment_level, age_grp) %>%
    summarise( yes = sum(yes), no = sum(no)) %>%
    transmute( hearing_impairment_level, age_grp, sex = 'both', yes, no)
) %>% mutate(
  pct = ifelse( yes+no== 0, 0, yes / (yes + no))
) 

prevalences <- bind_rows(
  depr_by_impairment %>%
  mutate( disease = 'depression'),
  anx_by_impairment %>%
  mutate( disease = 'anxiety') ) 

dfle <- prevalences %>% 
  transmute(disease, sex, hearing_impairment_level, age_grp, pix = pct) %>%
  arrange( disease, sex,hearing_impairment_level, desc(age_grp) ) %>% 
  na.omit() %>%
  left_join(br_ibge_censo_2010, by = c('age_grp','sex')) %>% 
  mutate( aux = (1-pix)*Lxn ) %>% 
  group_by(disease, sex, hearing_impairment_level) %>%
  mutate( 
    dfle = cumsum(aux)/lx, 
    age_grp = factor( age_grp, levels = AGE_LABELS),
    dfle_over_Ex = dfle / Ex
    ) %>% 
  select( -aux ) %>% 
  arrange( disease, sex, hearing_impairment_level, age_grp ) 


fig2 <- dfle %>% 
  filter(disease == 'depression') %>%
  filter(!age_grp %in% c('<1','1-4','5-9','10-14','85-89','90+')) %>% 
  mutate(
    age_grp = as.character(age_grp),
    age_grp = case_when(
      age_grp == '80-84' ~ '80+',
      TRUE ~ as.character(age_grp)
    ) 
  ) %>% 
  ggplot(
    aes(
    x = age_grp,
    y = dfle_over_Ex,
    shape = sex,
    group = hearing_impairment_level
    )
  ) +
  geom_ribbon(
    aes(
      ymin = dfle_over_Ex, 
      ymax = 1, 
      fill = "Gap"
      ),
    alpha = .4, 
    color = "white"
  ) +
  geom_hline(yintercept = 1, linetype = "dotted", color = "black") +
  geom_point(size = .8, shape = 4, stroke = 1.5) +
  labs(
    y = expression("DFLE / e"[x]),
    x = 'Age group',
    fill = expression("Proportional impact of disease on e"[x])
  )+
  theme_bw(base_size = 15) +
  theme(
    legend.position = "top", 
    axis.text.x = element_text(angle = 90, vjust = .5),
    panel.grid = element_blank()
    ) +
  facet_grid(hearing_impairment_level ~ sex) +
  scale_fill_manual(
    name = "", 
    values = "darkblue",
    labels = expression("Proportional impact of depression on e"[x])
  ) +
  scale_y_continuous(
    limits = c(0.7, 1),
    labels = scales::percent_format(scale = 100)
  )

ggsave(
  filename = 'fig02.png',
  plot = fig2,
  width = 10,
  height = 8
)

ggsave(
  filename = 'fig02.pdf',
  plot = fig2,
  width = 10,
  height = 8
)

ggsave(
  filename = 'fig02.svg',
  plot = fig2,
  width = 10,
  height = 8
)

fig2
  
```

## Figure 03 (Supplementary Material): Depression prevalence among users of Brazilian Sign Language by 5-year age group and sex, Brazil, 2019

```{r dfle_figure03_suppl_material, include=TRUE}
# among users of Libras
# ever diagnosed with depression by a physician (g058)
depr_libras <- dfle_pns %>% 
  filter( g05801 == '1', combined %in% c('2','3','4') ) %>% # only users of Libras with at least some hearing impairment
  group_by( age_grp, sex, diagnosed_depression) %>%
  summarise( n = sum( Freq ) ) %>% 
  filter( diagnosed_depression != ' ') %>%
  pivot_wider( id_cols = c(1,2), names_from = 3, values_from = n ) %>% 
  rename( yes = `1`, no = `2`)

depr_libras <- bind_rows(
  depr_libras,
  depr_libras %>%
    group_by(age_grp) %>%
    summarise( yes = sum(yes), no = sum(no)) %>%
    transmute( age_grp, sex = 'both', yes, no)
  ) %>%
  mutate( pct = ifelse( yes+no== 0, 0, yes / (yes + no)) ) 

fig3 <- depr_libras %>%
  filter(!age_grp %in% c('<1','1-4','5-9','10-14','85-89','90+')) %>% 
  mutate(
    age_grp = as.character(age_grp),
    age_grp = case_when(
      age_grp == '80-84' ~ '80+',
      TRUE ~ as.character(age_grp)
    ) 
  ) %>% 
  ggplot(
    aes(
      x = age_grp,
      y = pct,
      group = sex
      )
  ) +
  geom_col(fill = "darkblue", width = .8) +
  geom_text(
    aes(label = scales::percent(round(pct, 2))),
    size = 3.5,
    nudge_y = 1.5/15,
    fontface = "bold",
    show.legend = FALSE,
    angle = 0
  ) +
  labs(
    y = 'People diagnosed with depression \n by a physician'
    ,x = 'Age groups'
    ,group = ''
    ,colour = ''
    ,fill = ''
  )+
  theme_bw(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    panel.spacing.x = unit(1.5, "lines"),
    legend.position = 'none'
  ) +
  scale_color_brewer(palette = "Dark2") +
  scale_y_continuous(
    limits = c(0, 1.10),
    labels = scales::percent_format(scale = 100),
    expand = c(0, 0)
  )+
  facet_wrap(~ sex) + coord_flip()

ggsave(
  filename = 'fig03.png',
  plot = fig3,
  width = 9,
  height = 5
)

ggsave(
  filename = 'fig03.pdf',
  plot = fig3,
  width = 9,
  height = 5
)

ggsave(
  filename = 'fig03.svg',
  plot = fig3,
  width = 9,
  height = 5
)

fig3
```

## Figure 04 (Supplementary Material): Depression prevalence by 5-year age group, sex and knowledge of Brazilian Sign Language

```{r dfle_figure04_suppl_material, include=TRUE}
# Depression (diagnosed by a physician)
# Comparison between users and not users of Libras
depr_compr <- dfle_pns %>% 
  group_by( age_grp, sex, libras, diagnosed_depression) %>%
  summarise( n = sum( Freq ) ) %>% 
  filter( diagnosed_depression != ' ') %>%
  pivot_wider( id_cols = c(1,2,3), names_from = 4, values_from = n ) %>% 
  rename( yes = `1`, no = `2`)
                 
depr_compr <- bind_rows(
  depr_compr,
  depr_compr %>%
    group_by( age_grp, libras) %>%
    summarise( yes = sum(yes), no = sum(no)) %>%
    transmute( age_grp, sex = 'both', libras, yes, no)
) %>%               
  mutate( pct = ifelse( yes + no == 0, 0, yes / (yes + no)) )

fig4 <- depr_compr %>%
  filter(!age_grp %in% c('<1','1-4','5-9','10-14','85-89','90+')) %>% 
  mutate(
    age_grp = as.character(age_grp),
    age_grp = case_when(
      age_grp == '80-84' ~ '80+',
      TRUE ~ as.character(age_grp)
    ) 
  ) %>% 
  na.omit() %>%
  ggplot(
    aes(
      x = age_grp,
      y = pct,
      group = libras,
      colour = libras
    )
  )+
  geom_line(
    aes(group = age_grp),
    size = 1,
    color = "grey26",
    alpha = 1
  ) +
  geom_point(size = 4, alpha = .8) +
  geom_stripped_cols(color = "white") +
  labs(
    y = 'People diagnosed with depression by a physician'
    ,x = 'Age groups'
    ,group = 'Knows brazilian sign \n language (Libras)'
    ,colour = 'Knows brazilian sign \n language (Libras)'
  )+
  theme_bw(base_size = 13) +
  theme(
    legend.position = 'top',
    panel.grid = element_blank()
  )+
  scale_color_brewer(palette = "Dark2") +
  scale_y_continuous(labels =
  scales::percent_format(scale = 100)) +
  facet_wrap(~sex) +
  coord_flip()

ggsave(
  filename = 'fig04.png',
  plot = fig4,
  width = 7,
  height = 5
)

ggsave(
  filename = 'fig04.pdf',
  plot = fig4,
  width = 7,
  height = 5
)

ggsave(
  filename = 'fig04.svg',
  plot = fig4,
  width = 7,
  height = 5
)


fig4
```

