---
title: "Table2 -- Paper on Deaf Depression-free Life Expectancy"
author: "Rafael Rodrigues de Moraes"
date: "2024-11-10"
output:
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Packages

The following packages are needed to load the datasets (`arrow` and
`readr`), applyt the complex survey design (`survey`) to the 
National Health Survey (PNS2019) selected microdata and
prepare the data frames (`dplyr`, `tidyr`) for the analyses in the
paper.

```{r libraries, message=FALSE, warning=FALSE, include=TRUE}
library(readr)     # read_delim
library(tidyr)     # pivot_wider
library(dplyr)     # %>%, filter, select, 
```

# Importing the baseline dataset
```{r baseline_dataset_import, include=TRUE}
dfle_pns <-
  readRDS(file = '/home/rstudio/data/dfle_baseline_dataset.rds')
```

# Table 2: Libras users and prevalence of depression for respondents with age between 5 and 40 years, Brazil, 2019

```{r dfle_table02, include=TRUE}
summary_aux <- dfle_pns %>% 
  filter( c008 >= 5, c008 <= 40 ) %>%
  group_by(combined, hearing_impaired, libras, diagnosed_depression) %>%
  summarise( n = sum(Freq))
  
summary_aux %>%
  group_by(combined, hearing_impaired, libras) %>%
  summarise( n = sum(n)) %>%
  na.omit() %>% 
  pivot_wider(id_cols = c(1,2), names_from = 3, values_from=4) %>%
  mutate( pct_libras_users = yes / (yes+no)) %>% 
  select( -c(yes,no)) %>%
  left_join( summary_aux %>% 
               group_by(combined, hearing_impaired, diagnosed_depression) %>% 
               summarise( n = sum(n)) %>%
               na.omit() %>% 
               pivot_wider(id_cols = c(1,2), names_from = 3, values_from=4) %>% 
               mutate( pct_depression = `1` / (`1`+`2`)) %>% 
               select( -c(3:5)) 
             ) %>%
  left_join( summary_aux %>% 
               filter( libras == 'yes') %>% 
               group_by(combined, hearing_impaired, diagnosed_depression) %>% 
               summarise( n = sum(n)) %>% 
               na.omit() %>% 
               pivot_wider(id_cols = c(1,2), names_from = 3, values_from=4) %>% 
               mutate( pct_depression_among_libras_users = `1` / (`1`+`2`)) %>%
               select( -c(3:5)) 
             ) %>% 
  ungroup() %>%
  select( -combined ) %>%
  knitr::kable()
```

