---
title: "Survey Design Validation -- Deaf Depression-free Life Expectancy Paper"
author: "Rafael Rodrigues de Moraes"
date: "2024-11-10"
output:
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Packages

The following packages are needed to load the datasets (`arrow` and
`readr`), applyt the complex survey design (`survey`) to the 
National Health Survey (PNS2019) selected microdata and
prepare the data frames (`dplyr`, `tidyr`) for the analyses in the
paper.

```{r libraries, message=FALSE, warning=FALSE, include=TRUE}
library(tidyverse)
library(arrow)     # read_parquet
library(survey)    # svydesign, postStratify, svytable
library(PNSIBGE)   # get_pns
library(dplyr)     # mutate, group_by, summarise, %>%
library(tidyr)     # pivot_wider

```

# Importing the processed datasets

## [`br_ibge_pns_2019`] National Health Survey 2019

```{r br_ibge_pns_2019, include=TRUE}
# 2019th edition from National Health Survey (PNS/IBGE)
br_ibge_pns_2019 <- read_parquet(
  file = '/home/rstudio/data/br_ibge_pns_2019/microdata.parquet'
) %>%
  mutate(across(everything(), as.numeric))
```

# Complex Survey Design

The National Health Survey (PNS) is a complex survey consisting of
strata, cluster and weights.

The same steps were followed in order to set up the survey design object
and post-stratify it as listed in the official R package `PNSIBGE`
(version 0.1.7).

```{r pns_survey_design, include=TRUE}
## Survey design (source: https://rdrr.io/cran/PNSIBGE/src/R/pns_design.R, accessed on September, 2nd 2022)
options(survey.lonely.psu="adjust")
options(survey.adjust.domain.lonely=TRUE)

pns2019_prior <- survey::svydesign(
  ids = ~upa_pns, 
  strata = ~v0024, 
  data = subset(br_ibge_pns_2019, !is.na(v0028)), 
  weights = ~v0028, 
  nest = TRUE
)
# Post-stratification
popc.types <- data.frame( 
  v00283=as.character(unique(br_ibge_pns_2019$v00283)), 
  Freq=as.numeric(unique(br_ibge_pns_2019$v00282))
)
popc.types <- popc.types[order(popc.types$v00283),]
pns2019_posterior <- survey::postStratify(
  design=pns2019_prior, 
  strata=~v00283, 
  population=popc.types
)
```

## Validation against processed microdata

Despite using exactly the same complex survey design, a full match with
the government official statistics was not entirely possible.

![PNS2019 - Validation with self-reported health
perception](./pns_validation.png){width="70%"}

```{r pns_validation, include=TRUE}
# validation against
#   https://sidra.ibge.gov.br/tabela/7666#/n1/all/v/2667/p/all/c2/all/c12258/all/d/v2667%200/l/v,p+c2,t+c12258/resultado
round( svytable(formula=~j001+c006, design=subset(pns2019_posterior, c008 >=18))/1e3, 0) %>%
  as.data.frame() %>%
  mutate(
    tab7666 = case_when(
      j001 %in% c(1,2) ~ 1,
      j001 == 3 ~ 2,
      j001 %in% c(4,5) ~ 3,
    )
    ,tab7666 = factor( tab7666, levels = 1:3, labels = c('Muito ruim e ruim','Regular','Muito bom e bom'))
    ,c006 = factor( c006, levels = 1:2, labels = c('Masculino', 'Feminino'))
  ) %>%
  na.omit() %>%
  group_by(tab7666, c006) %>%
  summarise( n = sum(Freq) ) %>%
  pivot_wider(id_cols = tab7666, names_from = c006, values_from = n)
```

```{r pns_validation_1, include=TRUE}
round( svytable(formula=~c006, design=subset(pns2019_posterior, c008 >= 18))/1e3, 0)
```

```{r pns_validation_2, include=TRUE}
round( sum(svytable(formula=~c006, design=subset(pns2019_posterior, c008 >= 18)))/1e3, 0)
```

*Conclusion:* although overall total and totals by sex are OK, among
health perception the numbers do not match. Unfortunately it was not
possible to reconcile the differences but as the total figures are the
same, we proceed with the analysis.

## Validation against official package `PNSIBGE`

```{r get_pns_data, include=TRUE}
pns2019 <- get_pns(year=2019)
```

```{r svydesign_validation_pnsibge, include=TRUE}
# reconciliation against
#   https://sidra.ibge.gov.br/tabela/7666#/n1/all/v/2667/p/all/c2/all/c12258/all/d/v2667%200/l/v,p+c2,t+c12258/resultado
round( svytable(formula=~J001+C006, design=subset(pns2019, C008 >=18))/1e3, 0) %>%
  as.data.frame() %>%
  mutate(
    tab7666 = case_when(
      grepl('bom',J001, ignore.case = TRUE) ~ 1,
      J001 == 'Regular' ~ 2,
      grepl('ruim',J001, ignore.case = TRUE) ~ 3,
    )
    ,tab7666 = factor( tab7666, levels = 1:3, labels = c('Muito ruim e ruim','Regular','Muito bom e bom'))
  ) %>%
  na.omit() %>%
  group_by(tab7666, C006) %>%
  summarise( n = sum(Freq) ) %>%
  pivot_wider(id_cols = tab7666, names_from = C006, values_from = n)
# # A tibble: 3 × 3
# # Groups:   tab7666 [3]
# tab7666             Homem Mulher
# <fct>               <dbl>  <dbl>
# 1 Muito ruim e ruim 53296  54574
# 2 Regular           17646  24349
# 3 Muito bom e bom    3610   5696
```

```{r svydesign_validation_pnsibge_total_by_sex, include=TRUE}
round( svytable(formula=~C006, design=subset(pns2019, C008 >= 18))/1e3, 0)
# C006
# Homem Mulher 
# 74553  84619
```

```{r svydesign_validation_pnsibge_overall_total, include=TRUE}
round( sum(svytable(formula=~C006, design=subset(pns2019, C008 >= 18)))/1e3, 0)
# [1] 159171
```
